---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file and run
     devtools::build_readme() to regenerate README.md. -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

# robotstxtr

<!-- badges: start -->
[![r-universe](https://bart-turczynski.r-universe.dev/robotstxtr/badges/version)](https://bart-turczynski.r-universe.dev/robotstxtr)
[![Pipeline](https://gitlab.com/bart-turczynski/robotstxtr/badges/main/pipeline.svg)](https://gitlab.com/bart-turczynski/robotstxtr/-/pipelines)
[![Coverage](https://gitlab.com/bart-turczynski/robotstxtr/badges/main/coverage.svg)](https://gitlab.com/bart-turczynski/robotstxtr/-/pipelines)
[![Docs](https://img.shields.io/website?url=https%3A%2F%2Fbart-turczynski.gitlab.io%2Frobotstxtr%2F&label=docs&logo=gitlab&logoColor=white&up_message=pkgdown&up_color=1f75cb)](https://bart-turczynski.gitlab.io/robotstxtr/)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![Project Status: Active](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
[![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.23018995.svg)](https://doi.org/10.5281/zenodo.23018995)
[![Zenodo](https://img.shields.io/badge/Zenodo-all_software-1682D4?logo=zenodo&logoColor=white)](https://zenodo.org/search?q=metadata.creators.person_or_org.identifiers.identifier:0000-0002-8788-7980)
[![OpenSSF Best Practices](https://www.bestpractices.dev/projects/15192/badge)](https://www.bestpractices.dev/projects/15192)
[![License](https://img.shields.io/gitlab/license/bart-turczynski%2Frobotstxtr)](https://gitlab.com/bart-turczynski/robotstxtr/-/blob/main/LICENSE.md)
[![Last commit](https://img.shields.io/gitlab/last-commit/bart-turczynski%2Frobotstxtr)](https://gitlab.com/bart-turczynski/robotstxtr/-/commits/main)
<!-- badges: end -->

Faithful `robots.txt` parsing and URL-allowed decisions for R, powered by
Google's own matcher.

`robotstxtr` evaluates whether a URL may be crawled according to a site's
`robots.txt` rules. It wraps a vendored, upstream-test-validated snapshot of
Google's open-source C++ `robots.txt` parser and matcher, so allow and disallow
decisions match Googlebot's behavior — including the quirky parts (longest-match
wins, `Allow`-beats-`Disallow` ties, `*`/`$` wildcards, case-sensitive paths,
percent-escape canonicalization). It can also fetch `robots.txt` for you under a
deterministic, conservative `httr2` policy, construct fetch origins with `rurl`,
and return richly typed, vectorized decisions with per-rule match metadata.

This is a faithful, upstream-test-validated replica of the pinned matcher, not a
claim of universal behavioral identity with Googlebot.

## Installation

Install the released version from r-universe:

```{r install}
install.packages(
  "robotstxtr",
  repos = c("https://bart-turczynski.r-universe.dev", "https://cloud.r-project.org")
)
```

The development version lives on GitLab. Its dependencies, including `rurl`,
are all on CRAN, so it installs with no extra repository configuration:

```{r install-dev}
# with pak
pak::pak("gitlab::bart-turczynski/robotstxtr")

# or with remotes
remotes::install_gitlab("bart-turczynski/robotstxtr")
```

Installing from source compiles the vendored C++ matcher and requires a C++17
toolchain.
Once dependencies have been obtained, installing, testing, and checking the
package does not make live network requests.

## Usage

Match one or more URLs against a `robots.txt` body you already have, fully
offline, with `allowed_by_robots_text()`:

```{r usage-text}
library(robotstxtr)

robots <- "user-agent: *
disallow: /private
allow: /private/public
"

x <- allowed_by_robots_text(
  robots,
  c(
    "https://example.com/index.html",     # not disallowed  -> TRUE
    "https://example.com/private/report", # under /private  -> FALSE
    "https://example.com/private/public"  # re-allowed      -> TRUE
  ),
  "mybot"
)

x$results[c("url", "allowed", "decision_source")]
#>                                  url allowed decision_source
#> 1     https://example.com/index.html    TRUE   default_allow
#> 2 https://example.com/private/report   FALSE   rule_disallow
#> 3 https://example.com/private/public    TRUE      rule_allow
```

`allowed` is the `TRUE`/`FALSE`/`NA` answer; `decision_source` explains why.

If you do not already have the `robots.txt` text, `allowed_by_robots_url()`
derives the governing origin, fetches it once per origin, and matches:

```{r usage-url}
allowed_by_robots_url("https://example.com/some/page", "mybot")
```

See the "Introduction to robotstxtr" vignette for the full walkthrough,
including match metadata, the fetch policy, and body inspection with
`robots_body()`.

## Validate a robots.txt document

Document validation is separate from URL allow/disallow evaluation. Validate
supplied text (or raw acquisition bytes) with `robots_validate_text()`:

```{r validate}
validation <- robots_validate_text(
  "user-agent: *\ndisallow: /private\nunknown-field: value\n"
)

validation$documents
validation$diagnostics
```

Use `robots_validate_url()` when the document must first be fetched. It calls
the same deterministic `robots_fetch()` path and validates each distinct stored
raw body once; acquisition limits and failures remain explicit evidence.

The profile is deliberately named `google-parser-compatible`. It describes the
pinned parser plus conservative structural checks, not universal crawler or
RFC 9309 validity. Validation reports file syntax and structure only: it does
not answer whether a particular URL is allowed, infer whether rules match the
webmaster's intent, or fetch and check remote sitemap resources.

## Engine-aware v1 contract

Multi-engine integrations should use the parallel versioned facade and select
both axes explicitly:

```{r engine-aware}
robots_evaluate_text_v1(
  robots,
  "https://example.com/private/report",
  robots_product_token = "Googlebot",
  robots_policy_ruleset = "google",
  matcher_backend = "google"
)
```

The result distinguishes neutral fetch evidence, engine policy, matcher
availability, and the final URL decision. Unsupported RFC 9309 and Bing
matchers return `capability_unavailable` rather than silently falling back to
Google.

The Yandex `matcher_backend` is available as of schema revision `2026-07-18.2`,
bounded to profile `yandex-0.1.0` and the `Yandex` and `YandexAdditionalBot`
crawlers (every other Yandex-backend token resolves to a checked
`unsupported_crawler` non-decision). It is an independent, unofficial
compatibility profile and makes no production-crawler parity claim. See
[the engine-aware contract](https://gitlab.com/bart-turczynski/robotstxtr/-/blob/main/design/engine-contract-v1.md) and
`robots_engine_contract_v1()` for schema revisions and capability metadata.

## Citation

If you use `robotstxtr` in your work, please cite it. Run
`citation("robotstxtr")` for the current citation, or see
[`CITATION.cff`](https://gitlab.com/bart-turczynski/robotstxtr/-/blob/main/CITATION.cff).

Each release is archived on Zenodo. Cite the concept DOI
[10.5281/zenodo.23018995](https://doi.org/10.5281/zenodo.23018995) to refer to
`robotstxtr` in general; it always resolves to the latest release. Use the
version-specific DOI shown on the [Zenodo
record](https://doi.org/10.5281/zenodo.23018995) for a particular release.
